Assessing the impacts of time to detection distribution assumptions on detection probability estimation

### 2017-05-18 Draft.tex
Fix remaining two comments to reviewers

Just as the exponential TTDD is a special case of the two-parameter gamma and Weibull TTDDs, so all four TTDDs in our analysis are special cases the three-parameter generalized gamma distribution.
A generalized gamma TTDD encompasses a diversity hazard functions \citep{Cox2007}, eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
However, maximum likelihood estimation of the generalized gamma has historically suffered from computational difficulties, unsatisifactory asymptotic normality at large sample sizes, and non-unique roots to the likelihood equation \citep{CoorayAnand2008, NoufailyJones2013}.
It may well be possible to implement our model with a generalized gamma TTDD, but especially when we consider the right-truncation of data from point-count surveys, we think model convergence would not be a trivial problem.

>>>>

Just as the exponential TTDD is a special case of the two-parameter gamma and Weibull TTDDs, so all four TTDDs in our analysis are special cases the three-parameter generalized gamma distribution.
A generalized gamma TTDD encompasses a diversity of hazard functions \citep{Cox2007}, 
eliminating the need to restrict analysis to lognormal, gamma, or Weibull TTDDs and the tail probabilities they imply.
However, maximum likelihood estimation of the generalized gamma has historically suffered from computational difficulties, unsatisifactory asymptotic normality at large sample sizes, and non-unique roots to the likelihood equation \citep{CoorayAnand2008, NoufailyJones2013}.
It may well be possible to implement our model with a generalized gamma TTDD, but especially when we consider the right-truncation of data from point-count surveys, we think model convergence would not be a trivial problem.

### 2017-05-18 ReviewerComments.tex
Fix remaining two comments to reviewers

Introduction\\
Line 36-41: I find mention of survival analyses here and later in the manuscript confusing for most readers - especially Ecologists, presuming this is the intended audience.  It may be due to the unorthodox use of time-to-detection instead of traditional time-to-event analyses often used to estimate survival.  I suggest either omitting the reference to survival analysis here and elsewhere in the manuscript, or using the time-to-event terminology in this sentence and providing citations for the cdf and pdfs outlined here for reference.  Then explain that within the manuscript you will subsequently refer to it as TTDD.  
\begin{quote}
We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
\jarad{We probably need to add more, e.g. intended audience. Philip?} 
\end{quote}

@@

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.
This manuscript applies to removal-only analyses.
Still, the concept can be carried over to removal-distance analyses, as we describe in a paragraph of the Discussion.
The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
This is ongoing work for us.
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
\adam{I do not deem this necessary.  We have not explicitly studied sample size.  The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
\jarad{I'll have to re-read, but overall providing a summary for the reader
seems like a good idea.}
\end{quote}

>>>>

Introduction\\
Line 36-41: I find mention of survival analyses here and later in the manuscript confusing for most readers - especially Ecologists, presuming this is the intended audience.  It may be due to the unorthodox use of time-to-detection instead of traditional time-to-event analyses often used to estimate survival.  I suggest either omitting the reference to survival analysis here and elsewhere in the manuscript, or using the time-to-event terminology in this sentence and providing citations for the cdf and pdfs outlined here for reference.  Then explain that within the manuscript you will subsequently refer to it as TTDD.  
\begin{quote}
Since we and the AE feel this comparison is useful, 
we have retained the comparison. 
But we have tried to make the equivalence more explicit: ``We analyze times to first 
detection as time-to-event data, as is done in parametric survival analysis, ..."
\end{quote}

@@

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.
This manuscript applies to removal-only analyses.
Still, the concept can be carried over to removal-distance analyses, as we describe in a paragraph of the Discussion.
The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
This is ongoing work for us.
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
We have attempted to address these comments in the discussion:
``If the estimation of abundance is the primary interest, 
then the choice of TTDD has large consequences ... we would be skeptical of 
abundance estimates that indicate a nonpeaked TTDD with median posterior 
estimates of $\pdet$ below 0.75 ...
An assumption of constant detection limits the shape of $\ftdt$ and 
thereby constrains uncertainty but at the cost of potentially sizable bias...
all found constant-detection models to be flawed, 
producing underestimates of abundance and too-narrow error estimates; 
these resulted in inadequate coverage and also overstatement of effect 
significance.''

We have also included computation times in Section \ref{M-sec:simfull}: 
``Average times for exponential, lognormal, Weibull, and gamma model fits were 
2.6, 5.1, 5.3, and 25+ hours [when mixed effects were included in both abundance 
and detection].'' Since these types of surveys are typically conducted yearly, 
we do not feel even a one day runtime is arduous.
\end{quote}
